---
title: "Homework 10"
author: Xiao Ming
---
### Suggested Solutions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) 
library(quantmod) # only need to do this once
```

# Part III
## Problem 1
### 1.1
```{r, message=FALSE, warning=FALSE} 

# note: plz use "message=FALSE, warning=FALSE" in the above ```{r, ...}
# these are called "chunk options" and can be used to get rid of warnings and messages in your output

billion = 1e9
t = "1970-01-01/2015-01-01"

china = getSymbols('MKTGDPCNA646NWDB',src='FRED',auto.assign=FALSE)
china = china[t]/billion

plot(china,main="Nominal Chinese GDP (in Billions of Dollars)")
```

## Problem 5
### 5.1
```{r, message=FALSE, warning=FALSE}
t = "2003-01-01/2015-12-31"

rc = getSymbols('PCECC96',src='FRED',auto.assign=FALSE)
rc = rc[t]
rc = rc/coredata(rc)[1]

plot(rc,main="Real Consumption Index (2003.Q1 = 1)")
```

## Part IV
## Problem 1
### 1.2
```{r, message=FALSE, warning=FALSE}
t = "1950-01-01/2015-10-01"

gdpQ = getSymbols('GDP',src='FRED',auto.assign=FALSE)
gdpQ = gdpQ[t]

# convert quarterly data into annual data
gdp = apply.yearly(gdpQ,mean) 

plot(gdp,main="U.S. Nominal GDP (in Billions of Dollars)")
```

### 1.3
```{r, message=FALSE, warning=FALSE}
x <- xts(rnorm(10),as.Date(1:10,origin="2000-01-01")) # here we randomly generate some x. You need to replace x with the appropriate variable for computing gdp growth rates here. 
gx = (x/lag(x)-1)*100 
gx = gx[-1,] #get rid of the first row, which is NA
barplot(gx,main="x Growth Rate")
```

## Problem 2
### 2.1
```{r, message=FALSE, warning=FALSE}
t = "2000-01-01/2014-12-01"

china = getSymbols('CHNCPIALLMINMEI',src='FRED',auto.assign=FALSE)
china = china[t]

# calculate yoy inflation rate
pi.china = (china/lag(china,12)-1)*100
pi.china = pi.china[-c(1:12),] #get rid of the first 12 rows, which are NA

plot(pi.china,main="Chinese Inflation")
```

### 2.2
```{r, message=FALSE, warning=FALSE}
# Suppose we want to know how many months Chinese inflation is above 2 percent
sum(pi.china > 2)
```

## Problem 3
### 3.8
```{r, message=FALSE, warning=FALSE}
# Suppose we want to average Chinese monthly yoy (year-on-year) inflation in 2010:
mean(pi.china["2010"])
```