---
title: "Homework 11"
author: Xiao Ming
---
### Suggested Solutions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) 
library(quantmod) # only need to do this once
```

# Part I
## Problem 1
### 1.1
```{r, message=FALSE, warning=FALSE} 

# note: plz use "message=FALSE, warning=FALSE" in the above ```{r, ...}
# these are called "chunk options" and can be used to get rid of warnings and messages in your output

billion = 1e9
t = "1970-01-01/2020-01-01"

us = getSymbols('GDPA',src='FRED',auto.assign=FALSE)
china = getSymbols('MKTGDPCNA646NWDB',src='FRED',auto.assign=FALSE)

us = us[t]
china = china[t]/billion

GDP = cbind(us,china)
colnames(GDP) <- c("US","China")

plot(GDP,legend.loc='topleft',auto.legend=TRUE,main="Nominal GDP (in Billions of Dollars)")
```

## Problem 3
### 3.1
```{r, message=FALSE, warning=FALSE}
t = "2005-01-01/2020-12-31"

con = getSymbols('USCONSTRQGSP',src='FRED',auto.assign=FALSE)
man = getSymbols('USMANRQGSP',src='FRED',auto.assign=FALSE)

con = con[t]
man = man[t]

con = con/coredata(con)[1]
man = man/coredata(man)[1]

data = cbind(con,man)
colnames(data) <- c("Construction","Manufacturing")
plot(data,legend.loc='topleft',auto.legend=TRUE,main="GDP by Sector")
```

## Part II
## Problem 1
### 1.2
```{r, message=FALSE, warning=FALSE}
t = "1950-01-01/2020-10-01"

gdpQ = getSymbols('GDP',src='FRED',auto.assign=FALSE)
gdpQ = gdpQ[t]

# convert quarterly data into annual data
gdp = apply.yearly(gdpQ,mean) 

plot(gdp,main="U.S. Nominal GDP (in Billions of Dollars)")
```

### 1.3
```{r, message=FALSE, warning=FALSE}
x = (gdp/lag(gdp)-1)*100
x = x[-1,] #get rid of the first row, which is NA
barplot(x,main="U.S. GDP Growth Rate based on Nomial GDP (Incorrect)")
```

## Problem 2
### 2.1
```{r, message=FALSE, warning=FALSE}
t = "2000-01-01/2020-12-01"

china = getSymbols('CHNCPIALLMINMEI',src='FRED',auto.assign=FALSE)
china = china[t]

# calculate yoy inflation rate
pi.china = (china/lag(china,12)-1)*100
pi.china = pi.china[-c(1:12),] #get rid of the first 12 rows, which are NA

plot(pi.china,main="Chinese Inflation")
```

### 2.2
```{r, message=FALSE, warning=FALSE}
# Suppose we want to know how many months Chinese inflation is above 2 percent
sum(pi.china > 2)
```
