---
title: "An Introduction to Connected Scatter Plots"
author: "J. Mao \n \n [https://jiamingmao.github.io/principles-of-economics](https://jiamingmao.github.io/principles-of-economics)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

Let's generate some data:

```{r}
t = seq(-10,10,.2)
x = 16*(sin(t))^3
y = 13*cos(t) - 5*cos(2*t) - 2*cos(3*t) - cos(4*t)
```

First, let's plot x and y using a standard scatter plot. Here's what it looks like:

```{r,fig.asp=1,out.width="60%"}
plot(x,y)
```

To produce a *connected* scatter plot, there are a couple ways. First:

```{r,fig.asp=1,out.width="60%"}
plot(x,y,type="o")
```

Now, you can change the point symbol "o" or the color of the lines. Here's how to do it:

```{r,fig.asp=1,out.width="60%"}
# pch controls the plot symbol. 
# See here for a complete list of symbols: http://www.endmemo.com/program/R/pchsymbols.php
plot(x,y,type="o",pch=17,col="pink") 
```

But what if you want different colors for points and lines? You can do this:

```{r,fig.asp=1,out.width="60%"}
plot(x,y,pch=17,col="purple")
lines(x,y,lwd=2,col="pink") # lwd controls line width
```

or this: (reversing the order)

```{r,fig.asp=1,out.width="60%"}
plot(x,y,lwd=2,type="l",col="pink")
points(x,y,pch=17,col="purple")
```

Now let's try some new data:

```{r}
t = 2001:2010
x = 1:10
y = runif(10)
```

Let's plot x and y using a connected scatter plot, but using arrows to go from one point to the next:

```{r,fig.asp=.8,out.width="75%"}
plot(x,y,pch=16,col="orange")
s = 1:(length(x)-1)
arrows(x[s], y[s], x[s+1], y[s+1], col="steelblue")
```

We can also add text labels on the scatter points. Let's label each point's t value:

```{r,fig.asp=.8,out.width="75%"}
plot(x,y,pch=16,col="orange")
text(x,y,labels=t,pos=4) #pos=1,2,3,4 adjusts position of text label
s = 1:(length(x)-1)
arrows(x[s], y[s], x[s+1], y[s+1], col="steelblue")
```

Finally, let's try some fancy ggplot themes:

```{r,message=FALSE,warning=FALSE}
require(ggplot2)
require(ggthemes)
data <- data.frame(t=t,x=x,y=y)
ggplot(data,aes(x=x,y=y,label=t)) +
  geom_point(col="orange",size=2) + 
  geom_text(fontface="bold",nudge_x=.4) +
  geom_segment(color="steelblue", 
               aes(xend=c(tail(x, n=-1), NA), yend=c(tail(y, n=-1), NA)),
               arrow=arrow()) +
  theme_wsj() + scale_colour_wsj()
```
